---
---

<B>
<FONT SIZE = 4, color = "blue">
<center><BR>
M5 ASSIGNMENT - Nonparametric Statistical Methods and Sampling
<BR>
<FONT SIZE = 3, color = "black">
ALY6015: INTERMEDIATE ANALYTICS
<BR>
<FONT SIZE = 2.5, color = "black">
NORTHEASTERN UNIVERSITY
<BR>
<FONT SIZE = 2.5, color = "black">
</B>
Student - Jayakumar Moris Udayakumar
<BR>
<FONT SIZE = 2.5, color = "black">
Instructor Name: Prof. Zhi (Richard) He
<BR>
<FONT SIZE = 2.5, color = "black">
Date: 24th Mar 2024
<BR></center></B>

<B>
<FONT SIZE = 2, color = "black">
Section 13-2: Traditional Hypothesis Testing
</B><BR>

Game Attendance

```{r }

#H0: median  = 3000
#H1: median != 3000

#data given
attendance = c(6210,3150,2700,3012,4875,3540,6127,2581,2642,2573,2792,2800,2500,3700,6030,5437,2758,3490,2851,2720)
alpha = 0.05
n1 = length(attendance)
game_med0 = 3000

#sample median
samplmed_game = median(attendance)

#std error
stderr_game = sd(attendance) / sqrt(n1)

#test value
t_stat_game = (samplmed_game - game_med0) / stderr_game

# Print the test value
cat("Test value:", t_stat_game)

#critical value
CVleft_game = qt((alpha/2), 19)
CVright_game = qt((1-alpha/2), 19)

cat("Critical value (left-tailed):", CVleft_game, "\n")
cat("Critical value (right-tailed):", CVright_game)

#Compare the absolute test statistic with the critical values
if (abs(t_stat_game) > CVright_game || abs(t_stat_game) < CVleft_game) {
  print("Reject null hypothesis: Population median is not equal to 3000")
} else {
  print("Fail to reject null hypothesis: There is no sufficient evidence to prove median is not equal to 3000")
}

```

Lottery Ticket Sales

```{r}

#H0: median = 200
#H1: median < 200

#data given
alpha = 0.05
n2 = 40

n_over200 = 25
n_less200 = 15

# Number of days sold fewer than 200 tickets
teststat_tickets <- binom.test(x = c(n_over200, n_less200), alternative = "greater")

testvalue_tickets <- teststat_tickets$statistic

# Print the test value
cat("Test value:", testvalue_tickets, "; ")

# Find critical value
cv_tkts <- qbinom(alpha, n2, 0.5)

# Print the critical value
cat("Critical value:", cv_tkts)

#Compare the absolute test statistic with the critical values
if (testvalue_tickets <= cv_tkts) {
  print("; Reject null hypothesis: median is less than 200")
} else {
  print("Fail to reject null hypothesis: There is no sufficient evidence to prove median is less than 200")
}

```


<B>
<FONT SIZE = 2, color = "black">
Section 13-3: Wilcoxon rank sum test
</B><BR>

Lengths of Prison Sentences

```{r warning=FALSE}

#H0: There is no difference in the sentence received by each gender
#H1: There is a difference in the sentence received by each gender

alpha = 0.05

male_gen <- c(8, 12, 6, 14, 22, 27, 32, 24, 26, 19, 15, 13)
female_gen <- c(7, 5, 2, 3, 21, 26, 30, 9, 4, 17, 23, 12, 11, 16)

# Perform two-sample t-test
wilcoxranksum_test_result <- wilcox.test(male_gen, female_gen, alternative = "two.sided")

# Get test statistic (t-value)
W_value_gen <- wilcoxranksum_test_result$statistic

# Print the test value
cat("Test value:", W_value_gen, "; ")


# Determine sample sizes
n_mal <- length(male_gen)
n_fem <- length(female_gen)

#critical value for a two-tailed test - Wilcoxon Rank Sum Table
cv_gend <- 123

# Print the critical value
cat("Critical value:", cv_gend)

# Make decision
if (W_value_gen >= cv_gend) {
  cat("; Reject the null hypothesis: There is a difference in the sentence received by each gender\n")
} else {
  cat("; ", "Fail to reject the null hypothesis: There is no sufficient evidence to prove that there is a difference in the sentence received by each gender\n")
}
```

Winning Baseball Games

```{r warning=FALSE}

#H0: There is no difference in the number of wins by NL and AL
#H1: There is a difference in the number of wins by NL and AL

alpha = 0.05

nat_leg = c(89, 96, 88, 101, 90, 91, 92, 96, 108, 100, 95)
amer_leg = c(108, 86, 91, 97, 100, 102, 95, 104, 95, 89, 88, 101)

#Perform two-sample t-test
w_test_baseball <- wilcox.test(nat_leg, amer_leg, alternative = "two.sided")

# Get test statistic (t-value)
wtestvalue_bb <- w_test_baseball$statistic

# Print the test value
cat("Test value:", wtestvalue_bb, "; ")

# Determine sample sizes
n_NL <- length(nat_leg)
n_AL <- length(amer_leg)

#critical value for a two-tailed test - Wilcoxon Rank Sum Table
cv_league <- 99

# Print the critical value
cat("Critical value:", cv_league)

# Make decision
if (abs(wtestvalue_bb) >= abs(cv_league)) {
  cat("; Reject the null hypothesis: There is a difference in the number of wins by NL and AL.\n")
} else {
  cat("; Fail to reject the null hypothesis: There is no sufficient evidence to prove that there is a difference in the number of wins by NL and AL\n")
}

```

<B>
<FONT SIZE = 2, color = "black">
Section 13-4: Table K based problem-solving
</B><BR>

```{r}

#q5_two-tailed
ws5 = 13
n5 = 15
alpha5 = 0.01
table_k5 = 16

#q6_one-tailed
ws6 = 32
n6 = 28
alpha6 = 0.025
table_k6 = 117

#q7_one-tailed
ws7 = 65
n7 = 20
alpha7 = 0.05
table_k7 = 60

#q8_two-tailed
ws8 = 22
n8 = 14
alpha8 = 0.10
table_k8 = 26

if (ws5 <= table_k5) 
{cat("Q5: Reject the null hypothesis\n")} else {cat("Q5: Fail to reject the null hypothesis\n")}
if (ws6 <= table_k6) 
{cat("Q6: Reject the null hypothesis\n")} else {cat("Q6: Fail to reject the null hypothesis\n")}
if (ws7 <= table_k7) 
{cat("Q7: Reject the null hypothesis\n")} else {cat("Q7: Fail to reject the null hypothesis\n")}
if (ws8 <= table_k8) 
{cat("Q8: Reject the null hypothesis\n")} else {cat("Q8: Fail to reject the null hypothesis\n")}

```


<B>
<FONT SIZE = 2, color = "black">
Section 13-5: Kruskal-Wallis test
</B><BR>

Mathematics Literacy Scores

```{r}

#H0: No difference in means between the selected countries/regions
#H1: There is a significant difference in means between the selected countries/regions

alpha = 0.05

west_hemis = c(527, 406, 474, 381, 411)
europe_reg = c(520, 510, 513, 548, 496)
eastern_asia = c(523, 547, 547, 391, 549)

# Perform Kruskal-Wallis test
kruskaltest_mathlit <- kruskal.test(list(west_hemis, europe_reg, eastern_asia))

# Get the test statistic
kruskaltestvalue_mathlit <- kruskaltest_mathlit$statistic

# Print the test value
cat("Test value:", kruskaltestvalue_mathlit, "; ")

#degrees of freedom
df_math <- length(list(west_hemis, europe_reg, eastern_asia)) - 1

#critical value
cv_mathlit <- qchisq(1 - alpha, df_math)

# Print the critical value
cat("Critical value:", cv_mathlit)

# Compare test statistic with critical value
if (kruskaltestvalue_mathlit > cv_mathlit) {
  print("Reject null hypothesis: There is a significant difference in means between the selected countries/regions.")
} else {
  print("Fail to reject null hypothesis: No difference in means between the selected countries/regions.")
}

```


<B>
<FONT SIZE = 2, color = "black">
Section 13-6: Spearman rank correlation coefficient
</B><BR>

Subway and Commuter Rail Passengers

```{r}

#H0: There is no relationship between the variables
#H1: There is a relationship between the variables

city <- c(1, 2, 3, 4, 5, 6)
sub <- c(845, 494, 425, 313, 108, 41)
com_rail <- c(39, 291, 142, 103, 33, 38)


#spearman correlation coef
spearman_cor <- cor(sub, com_rail, method = "spearman")

# Print spearman correlation coef
cat("spearman correlation coef value:", spearman_cor, "; ")

#critical value as per Spearman rank correlation coefficient table
criticalvalue_pgrs = 0.886

# Print the critical value
cat("Critical value:", criticalvalue_pgrs)

# Make decision
if (abs(spearman_cor) > abs(criticalvalue_pgrs)) {
  cat("Reject the null hypothesis: There is a difference in the number of wins by NL and AL.\n")
} else {
  cat("; ", "Fail to reject the null hypothesis: There is no sufficient evidence to prove that there is a difference in the number of wins by NL and AL\n")
}

```


<B>
<FONT SIZE = 2, color = "black">
Section 14-3: Random Number Simulation: Seed 20353
</B><BR>

Rolling a Die
```{r}

# Set the random seed for reproducibility
set.seed(20353)

# Function to simulate rolling a die until all faces are seen
roll_dice_until_all_faces_seen <- function() {
  seen <- c()
  rolls <- 0
  while (length(unique(seen)) < 6) {
    roll <- sample(1:6, 1, replace = TRUE)
    seen <- c(seen, roll)
    rolls <- rolls + 1
  }
  return(rolls)
}

# Simulate the experiment 1000 times
num_rolls <- replicate(1000, roll_dice_until_all_faces_seen())

# Calculate the experimental average
experimental_avg <- round(mean(num_rolls), 1)

# Calculate the theoretical average using the harmonic number
theoretical_avg <- 6 * 2.45

# Print the results
cat("Experimental average number of tosses:", experimental_avg, "\n")
cat("Theoretical average number of tosses:", round(theoretical_avg, 1))

```


Clay Pigeon Shooting

Task 1

```{r}

# Set random seed
set.seed(20353)

# Define function to simulate a round of shooting
simulate_shooting <- function() {
  bob_hits <- FALSE
  alice_hits <- FALSE
  
  # Alice shoots first
  if (runif(1) <= 0.6) {
    alice_hits <- TRUE
  } else {
    # Bob shoots if Alice misses
    if (runif(1) <= 0.8) {
      bob_hits <- TRUE
    }
  }
  
  # Determine winner
  if (alice_hits) {
    return("Alice")
  } else if (bob_hits) {
    return("Bob")
  } else {
    return("Draw")
  }
}

# Simulate 5000 rounds of shooting
results <- replicate(5000, simulate_shooting())

# Calculate experimental probabilities
experimental_prob_alice_wins <- mean(results == "Alice")
experimental_prob_bob_wins <- mean(results == "Bob")

# Calculate theoretical probabilities
theoretical_prob_alice_wins <- 0.6  # Alice wins if she hits on the first shot
theoretical_prob_bob_wins <- 0.4 * 0.8  # Bob wins if Alice misses (0.4) and then Bob hits (0.8)

# Print results
cat("Experimental probability of Alice winning:", round(100 * experimental_prob_alice_wins), "%\n")
cat("Experimental probability of Bob winning:", round(100 * experimental_prob_bob_wins), "%\n\n")

cat("Theoretical probability of Alice winning:", round(100 * theoretical_prob_alice_wins, 2), "%\n")
cat("Theoretical probability of Bob winning:", round(100 * theoretical_prob_bob_wins, 2), "%\n")



```
Task 2

```{r}

# Define function to simulate shooting until a winner is determined and return the number of shots fired
simulate_shots_fired <- function() {
  shots_fired <- 0
  
  while (TRUE) {
    shots_fired <- shots_fired + 1
    
    if (runif(1) <= 0.6) {  # Alice hits
      return(shots_fired)
    } else if (runif(1) <= 0.8) {  # Bob hits
      return(shots_fired)
    }
  }
}

# Simulate 5000 rounds of shooting and store the number of shots fired in each round
shots_fired <- replicate(5000, simulate_shots_fired())

# Calculate average number of shots fired
average_shots_fired <- mean(shots_fired)

# Print result
cat("Average number of shots fired:", round(average_shots_fired, 2))


```

